Reducing Acceptance Marks in Emerson-Lei Automata by QBF-solving

### 2022-04-24 paper/main.tex
progress in Experiments

We have used the tool \texttt{genltl} of the Spot
library~\cite{duret.16.atva2} to generate LTL formulas from
literature. The parametrized formulas were generated for all
combinations of parameter values from 1 to 5. Altogether, we have
generated 407 unique LTL formulas. These have been translated to
deterministic TELA by two state-of-the-art translators from the Owl
library~\cite{kretinsky.18.atva} and to nondeterministic TELA by
another two state-of-the-art translators, namely \texttt{ltl2tgba}
from the Spot library~\cite{duret.16.atva2}) and
\texttt{ltl3tela}~\cite{major.19.atva}. Each of the four automata sets
produced by individual tools contain less than 407 automata as not all
of the formulas are successfully translated due to timeout of 60
seconds. Further, we have removed automata with 0 or 1 acceptance
marks as there is little point reducing these. Table
\ref{tab:ltltranslators} shows the exact versions of used translators,
the homepage of the used tools and the number of automata produced by
the given tool from the LTL formulas after filtering.

@@

For each automata set (that is, the output of an individual LTL
translator), we have applied all single-level reductions and the
multi-level reduction, always with incremental approach and without
reduction of the number of clauses. The timeout for each QBF query was
set to 30 seconds. We have again filtered out the automata for which
any of the four \telatko runs timed out after 120 seconds. We do this
to make sure the numbers of acceptance marks can be meaningfully
compared between different reduction levels of the same automata set.

@@

\begin{table}[b]
\caption{LTL to automata translators used for the experimental evaluation
and the number of successfully translated formulas.}
\label{tab:ltltranslators}
\centering
\begin{tabular}{lr@{~~~}lr}
\toprule
tool        & version&homepage & automata \\
\midrule
\texttt{ltl2dela} & 21.0 & \multirow{2}{*}{\url{https://owl.model.in.tum.de/}} & 162\\
\texttt{ltl2dgra} & 21.0 & & 274\\
\texttt{ltl2tgba} & 2.10.4 & \url{https://spot.lrde.epita.fr} & 89\\
\texttt{ltl3tela} & 2.2.0  & \url{https://github.com/jurajmajor/ltl3tela} & 108\\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[h]
\caption{The cummulative number of acceptance marks before and after reduction
for various automata sets. \textit{X/Y solved aut.} means that out of
\textit{Y} successfully translated LTL formulas, every run of \telatko ended
within 120 seconds time limit on \textit{X} automata. We sum up the acceptance
marks only on these automata. The columns \textit{solver response} and
\textit{solver timeout} refer to last call to the QBF solver (that is, the
numbers of its successful responses within 30 seconds limit and timeouts).
}
\label{tab:cumm-data}
\centering
\input{table-data.tex}
\end{table}

Table~\ref{tab:heatmap} shows the number of automata coming from all
four sets that have a given number of original acceptance marks and a
given number of acceptance marks after multi-level reduction. The
table indicates that in many cases only 1 or 2 marks can be
saved. However, the achieved reduction is substantial for some
automata with a higher number of original acceptance marks. For
example, in 17 cases, we have successfully reduced 8 or more
acceptance marks to only 4. 
% Moreover, there were many timeouts of the last QBF query for manof these automata.
% Let us note that some that some reductions are the results are promising, e.g. in 17 cases, we have successfully
% reduced 8 or more acceptance marks to only 4.

>>>>

We have used the tool \texttt{genltl} of the Spot
library~\cite{duret.16.atva2} to generate LTL formulas from
literature. The parametrized formulas were generated for all
combinations of parameter values from 1 to 4. Altogether, we have
generated 348 unique LTL formulas. These have been translated to
deterministic TELA by two state-of-the-art translators from the Owl
library~\cite{kretinsky.18.atva} and to nondeterministic TELA by
another two state-of-the-art translators, namely \texttt{ltl2tgba}
from the Spot library~\cite{duret.16.atva2} and
\texttt{ltl3tela}~\cite{major.19.atva}. Each of the four automata sets
produced by individual tools contain less than 348 automata as not all
of the formulas are successfully translated due to timeout of 60
seconds. Further, we have removed automata with 0 or 1 acceptance
marks as there is little point reducing these. Table
\ref{tab:ltltranslators} shows the exact versions of used translators,
the homepage of the used tools and the number of automata produced by
the given tool from the LTL formulas after filtering.

@@

For each automata set (that is, the output of an individual LTL
translator), we have applied all single-level reductions and the
multi-level reduction, always with incremental approach and without
reduction of the number of clauses. The timeout for each QBF query was
set to 30 seconds.
%We have again filtered out the automata for which
%any of the four \telatko runs timed out after 120 seconds. We do this
%to make sure the numbers of acceptance marks can be meaningfully
%compared between different reduction levels of the same automata set.
Additionally, we have filtered one automaton from the \texttt{ltl3tela} set
that cannot be processed by \telatko due to an extensive use of system
resources.

@@

\begin{table}[b]
\caption{LTL to automata translators used for the experimental evaluation
and the number of successfully translated formulas.}
\label{tab:ltltranslators}
\centering
\begin{tabular}{lr@{~~~}lr}
\toprule
tool        & version&homepage & automata \\
\midrule
\texttt{ltl2dela} & 21.0 & \multirow{2}{*}{\url{https://owl.model.in.tum.de/}} & 129\\
\texttt{ltl2dgra} & 21.0 & & 234\\
\texttt{ltl2tgba} & 2.10.4 & \url{https://spot.lrde.epita.fr} & 70\\
\texttt{ltl3tela} & 2.2.0  & \url{https://github.com/jurajmajor/ltl3tela} & 90\\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[h]
\caption{The cummulative number of acceptance marks before and after reduction
for various automata sets.
%\textit{X/Y solved aut.} means that out of
%\textit{Y} successfully translated LTL formulas, every run of \telatko ended
%within 120 seconds time limit on \textit{X} automata. We sum up the acceptance
%marks only on these automata.
The columns \textit{solver response} and
\textit{solver timeout} refer to last call to the QBF solver (that is, the
numbers of its successful responses within 30 seconds limit and timeouts).
}
\label{tab:cumm-data}
\centering
\input{table-data.tex}
\end{table}

Table~\ref{tab:heatmap} shows the number of automata coming from all
four sets that have a given number of original acceptance marks and a
given number of acceptance marks after multi-level reduction. The
table indicates that in many cases only 1 or 2 marks can be
saved. However, the achieved reduction is substantial for some
automata with a higher number of original acceptance marks. For
example, in 22 cases, we have successfully reduced 7 or more
acceptance marks to only 4.
% Moreover, there were many timeouts of the last QBF query for manof these automata.
% Let us note that some that some reductions are the results are promising, e.g. in 17 cases, we have successfully
% reduced 8 or more acceptance marks to only 4.
